---
title: "XAL2: Sácale jugo a tus datos"
subtitle: "Sesión 1"
author: "Said Muñoz, Miguel Nuñez"
date: "6 de Agosto 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
```
## Introducción a R

Algunos comandos básicos
## Suma
```{r}
1 + 1
```

## Multiplicación
```{r}
10 * 10
```
## Logaritmo
```{r}
log10(100)
```
## Print Text
```{r}
"Hello World"
'Hello World'
```
## Histograma básico
```{r}
hist(npk$yield)
```
## Asignación de una variable
```{r}
test <- 1

2->test2

test = 1

x <- 3
x + 5

x<-x+5
```
## Instalación de paquetes
### CRAN (static version)
```{r}
install.packages(c("aqp", "soilDB", "soilReports", "soiltexture"))

library(soiltexture)

help(package = "soiltexture")
```

## Vector
```{r}
x<-c(1,3)
```

## Funciones
```{r}
add_two_numbers <- function(num1, num2) {
  final=log(num1+num2)
  return(final)
}
add_two_numbers(4,5)
```
### Ejemplo de función
Can you write a function that calculates the mean of 3 numbers?
```{r}
Media3Numeros<-function(numero1,numero2,numero3=1000){
  Media<-(numero1+numero2+numero3)/3
  #Media<-mean(c(numero1,numero2,numero3))
  return(Media)
}
```

## Secuencias
```{r}
seq(0, 30) # This is the same as just `0:30`

x<-seq(0, 30, 2) # Every third number
```
## Función Módulo
```{r}
seq(2,20)%%2
```
## Vectores de Strings
```{r}
animals <- c('mouse', 'rat', 'dog')
animals # Characters
```
## Función class
Conocer el tipo del vector
```{r}
class(weight_g)
class(animals)
```
## Función structure
Conocer más del vector
```{r}
str(weight_g)
str(animals)
```
## Concatenar 
Agregar elementos a un vector
```{r}
weight_g <- c(weight_g, 90) # add to the end of the vector
weight_g <- c(30, weight_g) # add to the beginning of the vector
weight_g
```

### Ejercicio
Pregunta: ¿Qué tipo de vector será?
```{r}
num_char <- c(1, 2, 3, 'a')
num_logical <- c(1, 2, 3, TRUE)
char_logical <- c('a', 'b', 'c', TRUE)
char_logical <- c('a', 'b', 'c', TRUE)
tricky <- c(1, 2, 3, '4')


#class(num_char)
#class(num_logical)
#class(char_logical)
#class(tricky)
```

## Loops and vectorization
```{r}
v <- c(2, 4, 6)
for (num in v) {
  print(seq(1,20,by=num))
}

w <- 0
for (num in v) {
  w <- w + num
}
w
```

### Ejemplo
Ejemplo de una función que suma elementos en un vector
```{r}
my_sum <- function(input_vector) {
  vector_sum <- 0
  for (num in input_vector){
    vector_sum <- vector_sum + num
  }
  return(vector_sum)
}

my_sum(v)
```
## Diferencias entre un string y un booleano
```{r}
sum(c("TRUE"))
sum(c(TRUE))
```

## Paquete para leer tablas a dataframe
```{r}
install.packages('readr') 

library("readr")

surveys <- readr::read_csv('https://ndownloader.figshare.com/files/2292169')

## Descargar el archivo
#download.file("https://ndownloader.figshare.com/files/2292169",
#              "portal_data.csv") # Saves this name in the current directory

## Si el archivo estuviera de forma local se podría leer como: 
#surveys <- readr::read_csv('portal_data.csv')
```
## Dataframe
A data frame is the representation of data in the format of a 
table where the columns are vectors that all have the same length. 
Because the columns are vectors, they all contain the same type of 
data as we discussed before the break (e.g., characters, integers, factors). 
We can see this when inspecting the structure of a data frame with the function str()

### Dimensiones de un dataframe
```{r}
dim(surveys)
```
### Número de filas del dataframe
```{r}
nrow(surveys)
```
### Número de columnas del dataframe
```{r}
ncol(surveys)
```
### Mostrar algunos elementos del dataframe
```{r}
head(surveys)
tail(surveys)

names(surveys) 
colnames(surveys)

rownames(surveys)
str(surveys)

summary(surveys)
```

### Indexing
```{r}
surveys[1, 1]

surveys[1, 6] 

surveys[, 1]

surveys[1]

surveys[1:3, 7] # first three elements in the 7th

surveys[3, ]    # the 3rd element for all columns

surveys[1:6, ]  # equivalent to head(surveys)

surveys[,-1]    # All columns, except the first

surveys[-c(7:34786),]

surveys["species_id"]

surveys[, "species_id"]

head(surveys$species_id)          # Result is a vector
```

## While
```{r}
numero=1
while(numero<200){
  numero<-numero+1}

```